### 2019-12-15 ic-journal/sos.tex
Synchronized all changes from .tex to .htex; clean up all reviewed highlights (adding 3 new)

The CCS formalisation starts with type definitions for action,
relabeling and then the  processes. We use the type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}''
for all labels (i.e. visible actions), where the type variable $\beta$ corresponds
to $\mathscr{L}$ (the set of names for labels) \hl{mentioned} at the beginning
of Section~\ref{ss:ccs}. (Thus the
cardinality of ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}''  \hlD{depends on its type variable:
\mbox{when $\beta$ is finite or countable, ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' is countable.}})
All labels are divided into input and output ones.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' is defined by HOL's \texttt{Datatype} package in the
following syntax:
\begin{lstlisting}
Datatype: Label = name 'b | coname 'b
End
\end{lstlisting}
Intuitively, ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' turns names in $\mathscr{L}$ into \emph{input and output labels}.
For instance, if the type $\beta$ is instantiated to \texttt{string}
(the type of all ASCII strings), then the HOL terms \HOLinline{\HOLConst{name} \HOLStringLit{a}} and
\HOLinline{\HOLConst{coname} \HOLStringLit{b}} denote the input \emph{label} $a$ and output
\emph{label} $\overline{b}$, respectively.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Action}}'' is the
union of \hl{all visible actions (input and output labels) and the invisible action} $\tau$
(\texttt{tau}). For instace, the input \emph{action} $a$ and output
\emph{action} $\overline{b}$ of the type ``\HOLinline{\HOLTyOp{string} \HOLTyOp{Action}}''
are denoted by \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLStringLit{a}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLStringLit{b}\ensuremath{)}}, respectively. \hl{On the
other hand,} \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLNumLit{1}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLNumLit{2}\ensuremath{)}} could be actions of the type ``\HOLinline{\HOLTyOp{num} \HOLTyOp{Action}}'', where
\HOLinline{\HOLTyOp{num}} is the type of natural numbers in HOL.

The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'', accounting for all CCS terms,
has two type variables $\alpha$ and $\beta$ corresponding
to the set of agent variables $\mathscr{X}$
and the set of names $\mathscr{L}$,
respectively. (Indeed the CCS syntax in Section~\ref{ss:ccs} is parametric
w.r.t. the choice of these two sets.)
%
The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'' is defined inductively by the
\texttt{Datatype} package
(here ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Relabeling}}'' is the type of all relabeling
  functions; we have also formalised relabeling, though it is not
  discussed below):
\begin{lstlisting}
Datatype: CCS = nil
              | var 'a
              | prefix ('b Action) CCS
              | sum CCS CCS
              | par CCS CCS
              | restr (('b Label) set) CCS
              | relab CCS ('b Relabeling)
              | rec 'a CCS
End
\end{lstlisting}
The above \hl{definition allows us to write} terms like \texttt{nil} and
\texttt{sum P Q} in HOL4. Their correspondences with conventional CCS notations
in the literature are given in Table~\ref{tab:ccsoperator}, where
most CCS operators have also more readable \hl{abbreviated} forms,
either for end users or for \TeX{} outputs. (All
  formal theorems and definitions in this paper
  are generated from HOL4. Also, by default, all theorems are fully
  specialised with outermost universal quantifiers \hl{removed.})

@@

  \item
\emph{rules} is a conjunction of implications (each

@@

% \emph{rules} is a conjunction of implications that will be the
%    same as the input term. 

@@

For instance, \hlD{the following proposition
need the above \emph{cases} theorem}:

@@

\hl{In particular,} the SOS rule \texttt{REC} (Recursion)
 says that, if we substitute all occurrences of the variable $A$ in $P$ to
$(\recu A P)$ and the resulting process has a transition to $P'$
with an action $u$, then $(\recu A P)$ has the same
transition. Here ``\HOLinline{\ensuremath{[}\HOLConst{rec}\;\;\HOLFreeVar{X}\;\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\;\;\HOLFreeVar{E}}'' is an abbreviation
for ``\HOLinline{\HOLConst{CCS_Subst}\;\;\HOLFreeVar{E}\;\;\ensuremath{(}\HOLConst{rec}\;\;\HOLFreeVar{X}\;\;\HOLFreeVar{E}\ensuremath{)}\;\;\HOLFreeVar{X}}'', where \HOLinline{\HOLConst{CCS_Subst}} is a
recursive function substituting all \hl{occurrences} of a free variable with a CCS term.
For most CCS operators \HOLinline{\HOLConst{CCS_Subst}} just recursively goes into a deeper
level without changing anything, e.g.
\begin{alltt}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\ensuremath{\HOLFreeVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{2}}}\hfill{[CCS_Subst_sum]}
\end{alltt}
The only two insteresting cases are those for agent variables and recursion:
\begin{alltt}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLKeyword{if} \HOLFreeVar{Y} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{X} \HOLKeyword{then} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLKeyword{else} \HOLConst{var} \HOLFreeVar{Y}\hfill{[CCS_Subst_var]}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLKeyword{if} \HOLFreeVar{Y} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{X} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E} \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{Y} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_Subst_rec]}
\end{alltt}
Notice that, variable substitutions only affect free
variables. For instance, the variable $Y$ in ``\HOLinline{\HOLConst{rec}\!\;\HOLFreeVar{Y}\!\;\HOLFreeVar{E}}'' is
bound and therefore the substitution ignores it.

A useful facility exploiting the interplay
between HOL4 and Standard ML (\hl{which} follows an idea of Nesi~\cite{Nesi:1992ve})
 is a recursive ML function that takes a CCS process and returns a theorem indicating all
  direct transitions of the process. (If the input process is
 infinitely branching, the function will not terminate, however.)
For instance, we know that the process $(a.\nil | \bar{a}.\nil)$ has three
\hl{immediate derivatives given by the following transitions}:
$(a.\nil | \bar{a}.\nil) \overset{a}{\longrightarrow}
(\nil | \bar{a}.\nil)$, $(a.\nil | \bar{a}.\nil)
\overset{\bar{a}}{\longrightarrow} (a.\nil | \nil)$ and $(a.\nil | \bar{a}.\nil)
\overset{\tau}{\longrightarrow} (\nil | \nil)$.
To completely describe all possible transitions of the process, the
following two facts have to be proved: (1) there are indeed the three
  transitions mentioned above; (2) there is no other
  transition. For large CCS processes it is surprisingly tedious to manually derive
all the possible transitions and prove the non-existence of other transitions.
This shows the usefulness of appealing to an ML function
\texttt{CCS\_TRANS\_CONV} that is designed to automate the whole process.
For instance, taking the input $(a.\nil | \bar{a}.\nil)$ the function returns the
following theorem which describes all its one-step transitions:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLTokenTransBegin\HOLFreeVar{u}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}}
   \ensuremath{(}\HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\HOLTokenDisj{}} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenDisj{}}
   \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}
\end{alltt}

>>>>

The CCS formalisation starts with type definitions for action,
relabeling and then the  processes. We use the type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}''
for all labels (i.e. visible actions), where the type variable $\beta$ corresponds
to $\mathscr{L}$ (the set of names for labels) mentioned at the beginning
of Section~\ref{ss:ccs}. (Thus the
cardinality of ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' depends on its type variable:
when $\beta$ is finite or countable, ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' is countable.)
All labels are divided into input and output ones.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' is defined by HOL's \texttt{Datatype} package in the
following syntax:
\begin{lstlisting}
Datatype: Label = name 'b | coname 'b
End
\end{lstlisting}
Intuitively, ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Label}}'' turns names in $\mathscr{L}$ into \emph{input and output labels}.
For instance, if the type $\beta$ is instantiated to \texttt{string}
(the type of all ASCII strings), then the HOL terms \HOLinline{\HOLConst{name} \HOLStringLit{a}} and
\HOLinline{\HOLConst{coname} \HOLStringLit{b}} denote the input \emph{label} $a$ and output
\emph{label} $\overline{b}$, respectively.
The type ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Action}}'' is the
union of all visible actions (input and output labels) and the invisible action $\tau$
(\texttt{tau}). For instace, the input \emph{action} $a$ and output
\emph{action} $\overline{b}$ of the type ``\HOLinline{\HOLTyOp{string} \HOLTyOp{Action}}''
are denoted by \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLStringLit{a}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLStringLit{b}\ensuremath{)}}, respectively. On the
other hand, \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{name} \HOLNumLit{1}\ensuremath{)}} and \HOLinline{\HOLConst{label} \ensuremath{(}\HOLConst{coname} \HOLNumLit{2}\ensuremath{)}} could be actions of the type ``\HOLinline{\HOLTyOp{num} \HOLTyOp{Action}}'', where
\HOLinline{\HOLTyOp{num}} is the type of natural numbers in HOL.

The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'', accounting for all CCS terms,
has two type variables $\alpha$ and $\beta$ corresponding
to the set of agent variables $\mathscr{X}$
and the set of names $\mathscr{L}$,
respectively. (Indeed the CCS syntax in Section~\ref{ss:ccs} is parametric
w.r.t. the choice of these two sets.)
%
The type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS}}'' is defined inductively by the
\texttt{Datatype} package
(here ``\HOLinline{\ensuremath{\beta} \HOLTyOp{Relabeling}}'' is the type of all relabeling
  functions; we have also formalised relabeling, though it is not
  discussed below):
\begin{lstlisting}
Datatype: CCS = nil
              | var 'a
              | prefix ('b Action) CCS
              | sum CCS CCS
              | par CCS CCS
              | restr (('b Label) set) CCS
              | relab CCS ('b Relabeling)
              | rec 'a CCS
End
\end{lstlisting}
The above definition allows us to write terms like \texttt{nil} and
\texttt{sum P Q} in HOL4. Their correspondences with conventional CCS notations
in the literature are given in Table~\ref{tab:ccsoperator}, where
most CCS operators have also more readable abbreviated forms,
either for end users or for \TeX{} outputs. (All
  formal theorems and definitions in this paper
  are generated from HOL4. Also, by default, all theorems are fully
  specialised with outermost universal quantifiers removed.)

@@

  \item \emph{rules} is a conjunction of implications (each

@@

@@

For instance, the following proposition \hl{needs} the above \emph{cases} theorem:

@@

In particular, the SOS rule \texttt{REC} (Recursion)
 says that, if we substitute all occurrences of the variable $A$ in $P$ to
$(\recu A P)$ and the resulting process has a transition to $P'$
with an action $u$, then $(\recu A P)$ has the same
transition. Here ``\HOLinline{\ensuremath{[}\HOLConst{rec}\;\;\HOLFreeVar{X}\;\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\;\;\HOLFreeVar{E}}'' is an abbreviation
for ``\HOLinline{\HOLConst{CCS_Subst}\;\;\HOLFreeVar{E}\;\;\ensuremath{(}\HOLConst{rec}\;\;\HOLFreeVar{X}\;\;\HOLFreeVar{E}\ensuremath{)}\;\;\HOLFreeVar{X}}'', where \HOLinline{\HOLConst{CCS_Subst}} is a
recursive function substituting all occurrences of a free variable with a CCS term.
For most CCS operators \HOLinline{\HOLConst{CCS_Subst}} just recursively goes into a deeper
level without changing anything, e.g.
\begin{alltt}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\ensuremath{\HOLFreeVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{1}}} \HOLSymConst{\ensuremath{+}} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{\HOLFreeVar{E}\sb{\mathrm{2}}}\hfill{[CCS_Subst_sum]}
\end{alltt}
The only two insteresting cases are those for agent variables and recursion:
\begin{alltt}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLKeyword{if} \HOLFreeVar{Y} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{X} \HOLKeyword{then} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLKeyword{else} \HOLConst{var} \HOLFreeVar{Y}\hfill{[CCS_Subst_var]}
\HOLTokenTurnstile{} \ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \HOLKeyword{if} \HOLFreeVar{Y} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{X} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{Y} \HOLFreeVar{E} \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{Y} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLFreeVar{E}\sp{\prime}}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_Subst_rec]}
\end{alltt}
Notice that, variable substitutions only affect free
variables. For instance, the variable $Y$ in ``\HOLinline{\HOLConst{rec}\!\;\HOLFreeVar{Y}\!\;\HOLFreeVar{E}}'' is
bound and therefore the substitution ignores it.

A useful facility exploiting the interplay
between HOL4 and Standard ML (which follows an idea of Nesi~\cite{Nesi:1992ve})
 is a recursive ML function that takes a CCS process and returns a theorem indicating all
  direct transitions of the process. (If the input process is
 infinitely branching, the function will not terminate, however.)
For instance, we know that the process $(a.\nil | \bar{a}.\nil)$ has three
immediate derivatives given by the following transitions:
$(a.\nil | \bar{a}.\nil) \overset{a}{\longrightarrow}
(\nil | \bar{a}.\nil)$, $(a.\nil | \bar{a}.\nil)
\overset{\bar{a}}{\longrightarrow} (a.\nil | \nil)$ and $(a.\nil | \bar{a}.\nil)
\overset{\tau}{\longrightarrow} (\nil | \nil)$.
To completely describe all possible transitions of the process, the
following two facts have to be proved: (1) there are indeed the three
  transitions mentioned above; (2) there is no other
  transition. For large CCS processes it is surprisingly tedious to manually derive
all the possible transitions and prove the non-existence of other transitions.
This shows the usefulness of appealing to an ML function
\texttt{CCS\_TRANS\_CONV} that is designed to automate the whole process.
For instance, taking the input $(a.\nil | \bar{a}.\nil)$ the function returns the
following theorem which describes all its one-step transitions:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLTokenTransBegin\HOLFreeVar{u}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}}
   \ensuremath{(}\HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\HOLTokenDisj{}} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenDisj{}}
   \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}
\end{alltt}

### 2019-12-15 ic-journal/unique.tex
Synchronized all changes from .tex to .htex; clean up all reviewed highlights (adding 3 new)

\begin{proof}{(sketch from~\cite{sangiorgi2017equations})}
  Let $n$ be the length (i.e., the number of one-step transitions) of
a transition $\ct{\til P}\Arr\mu R$, and
let $\ctpp {\til P}$ and $\ctpp {\til Q}$  be the processes obtained
from  $\ct {\til P}$ and $\ct {\til Q}$ by unfolding the \hl{definition
of $\mcontrBIS$ for} $n$ times. Thus in $\qctpp$ each hole is
underneath at least $n$ prefixes, and \hl{therefore} cannot contribute to an action
in the first $n$ transitions. \hl{Moreover}, all \hl{involved} contexts use guarded sums.

@@

Theorem~\ref{t:contraBisimulationU} brings a new proof technique for
bisimilarity, which is  less restrictive than Milner's
Theorem~\ref{t:Mil89} (but with the additional costs of checking $\mcontrBIS$
in addition to $\wb$).
However, compared with Milner's Theorem~\ref{t:Mil89s1}, there remains
\hlD{the limitation on the need of 
guarded sums}. This is mainly due to the fact that $\wb$ is not a
 congruence and also $\mcontrBIS$ is not a precongruence.
Inspired by rooted bisimilarity, to eliminate the restriction
on guarded sums we refine the idea of contractions by moving
to \emph{rooted contractions}:

@@

The above definition was found with the help of interactive theorem proving.
The following two principles were adopted when \hl{manually} searching
for a possible definition: (1) the definition should not be coinductive,
along the lines of rooted bisimilarity $\rapprox$ (Def.~\ref{d:rootedBisimilarity});
(2) the definition should be built on top of  the existing \emph{contraction}
relation $\mcontrBIS$.
Furthermore, we needed to \hl{prove} that the definition \hl{being found}
indeed yields the coarsest precongruence
\hl{contained} in $\mcontrBIS$. The proof is similar with
the analogous result for $\rapprox$. See Section~\ref{s:coarsest} for
more details.

@@

For this new relation, the analogous of Lemma~\ref{l:uptocon} and of
Theorem~\ref{t:contraBisimulationU} can now be stated without
constraints on \hl{summation}.
The schema of the proofs is almost identical, because all 
properties of $\rcontr$ needed in this proof is its precongruence, which is
indeed true for \hl{all weakly guarded contexts:} % removed "direct sums"
\begin{lemma}
\label{l:ruptocon}
Let $\til P$ and $\til Q$ be solutions (for $\rcontr$)
 of a system of weakly guarded contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then there is a context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arr{\mu} R'
 \wb \ctp{\til Q}$ for some $R'$.
\end{lemma}

@@

\begin{theorem}[unique solution of rooted contractions]
\label{t:rcontraBisimulationU}
Let $E_i$ be weakly guarded with free variables \hlD{in} $\til X$,
and let $\til P \rcontr \til E\{\til P/\til X\}$,
    $\til Q \rcontr \til E\{\til Q/\til X\}$. Then $\til P \rapprox \til Q$.
\end{theorem}

>>>>

\begin{proof}{(sketch from~\cite{sangiorgi2017equations})}
  Let $n$ be the length (i.e., the number of one-step transitions) of
a transition $\ct{\til P}\Arr\mu R$, and
let $\ctpp {\til P}$ and $\ctpp {\til Q}$  be the processes obtained
from  $\ct {\til P}$ and $\ct {\til Q}$ by unfolding the definition
of $\mcontrBIS$ for $n$ times. Thus in $\qctpp$ each hole is
underneath at least $n$ prefixes, and therefore cannot contribute to an action
in the first $n$ transitions. Moreover, all involved contexts use guarded sums.

@@

Theorem~\ref{t:contraBisimulationU} brings a new proof technique for
bisimilarity, which is  less restrictive than Milner's
Theorem~\ref{t:Mil89} (but with the additional costs of checking $\mcontrBIS$
in addition to $\wb$).
However, compared with Milner's Theorem~\ref{t:Mil89s1}, there remains
the limitation on the need of 
guarded sums. This is mainly due to the fact that $\wb$ is not a
 congruence and also $\mcontrBIS$ is not a precongruence.
Inspired by rooted bisimilarity, to eliminate the restriction
on guarded sums we refine the idea of contractions by moving
to \emph{rooted contractions}:

@@

The above definition was found with the help of interactive theorem proving.
The following two principles were adopted when manually searching
for a possible definition: (1) the definition should not be coinductive,
along the lines of rooted bisimilarity $\rapprox$ (Def.~\ref{d:rootedBisimilarity});
(2) the definition should be built on top of  the existing \emph{contraction}
relation $\mcontrBIS$.
Furthermore, we needed to prove that the definition being found
indeed yields the coarsest precongruence
contained in $\mcontrBIS$. The proof is similar with
the analogous result for $\rapprox$. See Section~\ref{s:coarsest} for
more details.

@@

For this new relation, the analogous of Lemma~\ref{l:uptocon} and of
Theorem~\ref{t:contraBisimulationU} can now be stated without
constraints on summation.
The schema of the proofs is almost identical, because all 
properties of $\rcontr$ needed in this proof is its precongruence, which is
indeed true for all weakly guarded contexts: % removed "direct sums"
\begin{lemma}
\label{l:ruptocon}
Let $\til P$ and $\til Q$ be solutions (for $\rcontr$)
 of a system of weakly guarded contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then there is a context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arr{\mu} R'
 \wb \ctp{\til Q}$ for some $R'$.
\end{lemma}

@@

\begin{theorem}[unique solution of rooted contractions]
\label{t:rcontraBisimulationU}
Let $E_i$ be weakly guarded with free variables in $\til X$,
and let $\til P \rcontr \til E\{\til P/\til X\}$,
    $\til Q \rcontr \til E\{\til Q/\til X\}$. Then $\til P \rapprox \til Q$.
\end{theorem}
